### 2015-01-14 fundamentals/impl.tex
small changes

Many fuel cycle simulators rely on \gls{COTS} and Windows-only software that limits 
performance on and compatibility with resource computing infrastructures. This limitation severely 
increases the wall-clock time necessary to conduct parameterized sensitivity 
analyses and other multi-simulation studies. \Cyclus, on the other hand, is 
primarily written in \texttt{C++} and relies on 
libraries supported by Linux and UNIX (including Ubuntu and OSX) platforms, 
which are flexible and support parallelization. 
Furthermore, the core infrastructure and related archetypes are free and 
open source, BSD-3-clause licensed. No part of \Cyclus is proprietary or based 
on \gls{COTS}, Windows-only software. \Cyclus can therefore be easily deployed 
on large computer systems, such as \gls{HTC} systems.

@@

\subsubsection{Dynamically Loadable Libraries}
% (diagram)

A key innovation is that \Cyclus  implements this generic \gls{API} and modular 
architecture via a suite of dynamically loadable plug-in libraries.  Though 
common in modern software architecture, such a plug-in pattern has 
not previously been implemented in nuclear fuel cycle simulators.  
Dynamically loadable libraries are the primary mechanism for extending \Cyclus' capability. This approach provides encapsulation: the core of the code operates
independently from the individual plug-in libraries. Thus, any
customization or extension is implemented only in the loadable
library.

This approach encourages efficient, targeted contribution to the ecosystem of 
archetype libraries.  The 
researcher can focus on generating an archetype model within their
sphere of expertise, while relying on the contributions of others to fill 
in the other technologies in the simulation.  This strategy also allows individual developers to
explore different levels of complexity within their archetypes, including
wrapping other simulation tools as loadable libraries within \Cyclus.

@@

In large simulations many material objects may be changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features.
Immutability of compositions allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

>>>>

Groundbreaking insights and new classes of problems can be investigated by 
leveraging modern massive computing resources.
However, many fuel cycle simulators rely on \gls{COTS} and Windows-only software that limits 
performance on and compatibility with resource computing infrastructures. This limitation severely 
limits the possible scope of simulations and increases the wall-clock time necessary to conduct parameterized sensitivity 
analyses and other multi-simulation studies. \Cyclus, on the other hand, is 
primarily written in \texttt{C++} and relies on 
libraries supported by Linux and UNIX (including Ubuntu and OSX) platforms, 
which are flexible and support parallelization. 
Furthermore, the core infrastructure and related archetypes are free and 
open source, BSD-3-clause licensed. No part of \Cyclus is proprietary or based 
on \gls{COTS}, Windows-only software. \Cyclus can therefore be easily deployed 
on large computer systems, such as \gls{HTC} systems.

@@

\subsubsection{Dynamically Loadable Libraries}

The \Cyclus architecture encourages efficient, targeted contribution to the ecosystem of 
archetype libraries.
With \Cyclus, a researcher can focus on generating an archetype model within their
sphere of expertise, while relying on the contributions of others to fill 
in the other technologies in the simulation.  
Similarly, individual developers may explore different levels of complexity within their archetypes, including
wrapping other simulation tools as loadable libraries within \Cyclus.

\Cyclus acheives this behavior by implementing its generic \gls{API} and 
modular architecture via a suite of dynamically loadable plug-in libraries.  
Though common in modern software architecture, such a plug-in pattern has not 
previously been implemented in nuclear fuel cycle simulators.  Dynamically 
loadable libraries are the primary mechanism for extending \Cyclus' capability. 
This approach provides encapsulation: the core of the code operates 
independently from the individual plug-in libraries. Thus, any customization or 
extension is implemented only in the loadable library.

@@

In large simulations many material objects may be changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database size impacts,
compositions in \Cyclus have some special features.
Immutability of compositions allows multiple material objects to all hold
references to the same composition object safely.  When operations that do not 
alter the composition are performed on a material, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}
